---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align='center',
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# R package: ATE

The main aim of `ATE` is to provide a user-friendly interface for nonparametric efficient inference of average 
treatment effects for observational data. The package provides point estimates for average treatment 
effects, average treatment effect on the treated and can also handle the case of multiple treatments.
The package also allows inference by consistent variance estimates.

#### Requirements
- `R (>=3.2.0)`
- `Rcpp (>=0.12.0)`
- `RcppArmadillo`
- `Matrix`

## Installation

The package can be installed from CRAN:
```{r, eval = FALSE}
install.packages("ATE")
```

Alternatively, we can directly install from Github using the `devtools` package:

```{r, eval = FALSE}
devtools::install_github("asadharis/ATE")
```

## Key Features
* Ease of use: The main function `ATE` requires only a numeric matrix `X` of covariates, numeric vector `Y` of response 
and `treat` vector indicating treatment assignment.
```{r}
set.seed(1)
library(ATE)

#Generate some data
n <- 500
X1 <- matrix(rnorm(n * 5), ncol = 5)
X2 <- matrix(rbinom(3 * n, 1, 0.4), ncol = 3)
X <- cbind(X1, X2)
prop <- 1 / (1 + exp(X[, 1] - 0.5 * X[, 2] + 0.25 * X[, 3] + X[, 6] + 0.5 * X[, 8]))
treat <- rbinom(n, 1, prop)
Y <- 10 * treat + (2 * treat - 1) *
  (X[, 1] - 0.5 * X[, 2] + 0.25 * X[, 3] + X[, 6] + 0.5 * X[, 8]) + rnorm(n)

#Fit ATE object
fit1 <- ATE(Y, treat, X)
summary(fit1)
```

* `plot` function for demonstrating effect of covariate balancing for continuous and binary 
covariates.

```{r,fig.keep=c(1,6)}
plot(fit1)
```

* We can also estimate the average treatment effect on the treated. 

```{r}
fit2 <- ATE(Y, treat, X, ATT = TRUE)
summary(fit2)
```

* `ATE` automatically detects and estimates the case of multiple treatment arms. 

```{r}
treat <- rbinom(n, 3, prop)
Y <- 10 * treat + (2 * treat - 1) *
     (X[, 1] - 0.5 * X[, 2] + 0.25 * X[, 3] + X[, 6] + 0.5 * X[, 8]) +
     rnorm(n)
fit3 <- ATE(Y,treat,X)
summary(fit3)
```

```{r,fig.keep=c(1,6)}
plot(fit3)
```

* `ATE` uses the R packages `Rcpp` and `RcppArmadillo` to improve run-time. This allows us to handle big data efficiently.

Below we present the example for 10,000 observations and 800 covariates on an Intel® Core™ i5-3337U Processor.

```{r}
n <- 10000
X1 <- matrix(rnorm(n * 500), ncol = 500)
X2 <- matrix(rbinom(300 * n, 1, 0.4), ncol = 300)
X <- cbind(X1, X2)
prop <- 1 / (1 + exp( X[, 1] - 0.5 * X[, 2] + 0.25 * X[, 3] +  X[, 6] + 0.5 * X[, 8]))
treat <- rbinom(n, 1, prop)
Y <- 10 * treat + (2 * treat - 1) * 
  (X[, 1] - 0.5 * X[, 2] + 0.25 * X[, 3] + X[, 6] + 0.5 * X[, 8]) +
  rnorm(n)

system.time({
  fit4 <- ATE(Y, treat, X)
})
```

## Acknowledgements 

I would like to express my deep gratitude to Professor Gary Chan, my research supervisor, for his patient guidance, enthusiastic encouragement and useful critiques of this project.
